Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-05-22 ms.tex
Add explanation of detection process

Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$. For nonprecessing systems, $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies. In other words, $H_{\pi/2}(\omega) = i H_0(\omega)$ for all $\omega$. This linear combination isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform,
%
\begin{eqnarray}
	\fl H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
	\left(\begin{array}{cc}
		F_{+,i}(h_i, \delta) \\
		F_{\times,i}(h_i, \delta)
	\end{array}\right)^\intercal
	\left(\begin{array}{cc}
		\cos{2\psi} & -\sin{2\psi} \\
		\sin{2\psi} & \cos{2\psi}
	\end{array}\right)
	\left(\begin{array}{cc}
		\frac{1}{2}\left(1+\cos^2{\iota}\right) & 0 \\
		0 & \cos{\iota}
	\end{array}\right) \nonumber\\\fl &\cdot&
	\left(\begin{array}{cc}
		\cos{2\phi_c} & -\sin{2\phi_c} \\
		\sin{2\phi_c} & \cos{2\phi_c}
	\end{array}\right)
	\left(\begin{array}{cc}
		1 \\
		i
	\end{array}\right) H_0(\omega; \boldsymbol\theta_\mathrm{in}).
\end{eqnarray}
%
This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8). The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}. The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
%
\begin{equation}\label{eq:horizon}
D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
\end{equation}
%
More succinctly, we can write the signal received by detector $i$ in terms of an amplitude $\rho_i$, a phase $\gamma_i$, and a time delay $\tau_i$,
%
\begin{equation}\label{eq:signal-model}
	H_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} H_0(\omega; \boldsymbol\theta_\mathrm{in}).
\end{equation}

Detection of \acp{GW} from \ac{CBC} events consists of convolving the \dots explain matched filtering here \dots

\begin{eqnarray}
\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{H_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) X_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega
\end{eqnarray}

>>>>

Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$. For nonprecessing systems, $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies. In other words, $H_{\pi/2}(\omega) = i H_0(\omega)$ for all $\omega$. This linear combination isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform,
%
\begin{eqnarray}\label{eq:full-signal-model}
	\fl H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
	\left(\begin{array}{cc}
		F_{+,i}(h_i, \delta) \\
		F_{\times,i}(h_i, \delta)
	\end{array}\right)^\intercal
	\left(\begin{array}{cc}
		\cos{2\psi} & -\sin{2\psi} \\
		\sin{2\psi} & \cos{2\psi}
	\end{array}\right)
	\left(\begin{array}{cc}
		\frac{1}{2}\left(1+\cos^2{\iota}\right) & 0 \\
		0 & \cos{\iota}
	\end{array}\right) \nonumber\\\fl &\cdot&
	\left(\begin{array}{cc}
		\cos{2\phi_c} & -\sin{2\phi_c} \\
		\sin{2\phi_c} & \cos{2\phi_c}
	\end{array}\right)
	\left(\begin{array}{cc}
		1 \\
		i
	\end{array}\right) H_0(\omega; \boldsymbol\theta_\mathrm{in}).
\end{eqnarray}
%
This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8). The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}. The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
%
\begin{equation}\label{eq:horizon}
D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
\end{equation}
%
More succinctly, we can write the signal received by detector $i$ in terms of direct observables, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
%
\begin{equation}\label{eq:signal-model}
	H_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} H_0(\omega; \boldsymbol\theta_\mathrm{in}).
\end{equation}

The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from \Eref{eq:gaussian-likelihood} with the above signal model. This takes the form of a bank of matched filters, or template waveforms that are continually convolved with the incoming data stream. The matched filter or maximum likelihood estimates are given by
%
\begin{eqnarray}
\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{H_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) X_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
\end{eqnarray}
